---
title: "GDP and life expectancy"
output:
  html_document:
    fig_height: 3
    fig_width: 6
---

```{r setup, echo=FALSE}
# hide code and messages from output
# set echo = TRUE to show code in output
# set message = TRUE to show messages in output
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r load-packages}
library(ggplot2)
library(dplyr)
```

```{r load-dataset}
gapminder <- read.csv("data/gapminder.csv")
```

The Gapminder dataset contains life expectancy and GDP per capita information for countries between for the following years: `r unique(gapminder$year)`.

```{r set-year}
# change here for analysis for another year
# make sure to choose a year for which there is data available
analysis_year <- 2007
```

In this report we analyze the relationship between GDP and life expectancy in `r analysis_year` overall and by year.

```{r filter-for-year}
gap_subset <- gapminder %>%
  filter(year == analysis_year)
```

The figure below shows the relationship between these variables in all countries that we have data for in `r analysis_year`.

```{r overall}
ggplot(data = gap_subset, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  xlab("GDP per capita") +
  ylab("Life expectancy")
```

It might also be of interest to visualize this relationship conditional on continent.

```{r by-continent}
ggplot(data = gap_subset, aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_point() +
  xlab("GDP per capita") +
  ylab("Life expectancy")
```
